```{r setup, include=FALSE, warning=FALSE, message=FALSE}
library(data.table)
library(readxl)
library(lubridate)
library(ggplot2)
library(tidyr)
library("GGally") 
knitr::opts_chunk$set(echo = TRUE)
```

# Correlation Analysis on Measures of Credit/ Debit Cards 
### Sude Yagmur Ozturk

# Introduction

  Turkey, like other countries in the world, includes electronic payment systems in its banking and monetary structure. This new way of transferring money has also changed people's payment behavior, and a few terms came into our lives. Credit cards and debit cards are two of these terms. Credit/debit cards have some advantages comparing to cash. For example, cards provide more security than cash due to its traceability and password. Also, carrying these cards is more slighter than carrying all assets with cash. When adding banks' benefits in terms of commissions and government's benefits in terms of more taxable transactions into these user advantages, it was inevitable that credit/debit cards would be used more and more. 
  Debit cards do not have any advantages other than ease of use. Credit cards have remarkable  two advantages: Capability both to delay payments and to make installment plan for larger payments. Customers are provided financial opportunity to buy items that are too expensive to buy at once. These two features also bring money in banks, they can make interest income with customers' debts. Therefore, banks have started to release different types of credit cards in order to influence all customer segments. They are trying to specialize their credit card by extra campaigns and special installment plan offers. This competition is exaggerated especially between private banks, thus leading to the existence of cards such as gold, platinum, diamond. Their special offers can be regarded as not waiting in queue in airport or duplicating your car's oil for free etc.
  Easy payment, crazy installment plans and campaigns have merged with increase in shopping and consuming. These have accelerated the usage of credit/debit cards and it can be expected that their usage data show a trend line. However, consequences emerged afterwards. Individual credit card debits reached huge amounts in Turkey. Government forbade installment plans and late payments for some expensive products in 2014 to inhibit going into debt with credit cards. Media made propaganda to inform credit card customers about consequences. If these precautions answer the purpose, they may flatten the curve. That is why I chose three data sets about measures of credit/debit cards between April 2015 and September 2020. 
  There are 3 measures in data set below. The data are taken from Electronic Data Delivery System of Central Bank of Turkey: [EVDS](https://evds2.tcmb.gov.tr/index.php?/evds/serieMarket) .  First measure is monthly credit/debit cards transaction volume in thousand Turkish Liras in the second column. Second measure is monthly number of operations with credit/debit cards in the third column. Third measure is monthly credit card usage index in Turkey with respect to 04/2015 having 100 index point in the fourth column. First column is the date of 1st day of every month which represents the whole month. For example, 2018-03-01 involves the information about March 2018. This study covers data from April 2015 till September 2020 due to the availability of index measure. It is broad enough to study correlation analysis.
  Examining correlation between number of operations and transaction volume could be meaningful for studying whether change in transfer amounts depends on transaction numbers or not. Also, correlation information between the first two measures and usage index could provide insight about contribution of the first two measures to calculate the third one. 
  
# Manipulation of Data

  Before jumping into analysis, data must be imported from related excel file with read_excel. Some adjustments and manipulation of data are required for a meaningful output and easy forming of graphs. Data.table structure gives more opportunity for selecting and reshaping data. Also, Date format is more useful when subject is time series. Here are my code chunk to read  data with R, structure and first entries of data. For more information, [here](\fall20-sudeyagmur\files\hw2\HW2.Rmd) is my Rmd file with whole code chunks.

```{r, include= TRUE,echo = TRUE}
kredi_kartlari <- read_excel("C:/Users/Lenovo/Desktop/UZAKTAN BOGAZICI/IE360/kredi_kartlari2.xlsx")
credit_cards <- as.data.table(kredi_kartlari)
credit_cards[,Date := ymd(paste0(Date,"-01"))]
str(credit_cards)
head(credit_cards)
```

# Data Visualization and Analysis One by One

  First of all, visualizing the big picture must be done one by one in order to identify measures of interest. Creating histograms for each year is a good way to start analyzing behavior before studying correlation with pairs.
  
## Credit and Debit Cards' Transaction Volume

  These histograms below include density of transaction volume with credit/debit cards. Note that 2015 has 9 monthly values and 2020 has 10 monthly values. These numbers state transaction in thousands of Turkish Liras. There are 15 bins in the x axis.

```{r, echo=FALSE}
ggplot(data = credit_cards, aes(x= credit_transaction_vol))+
  geom_histogram(aes(y=..density..), color="aquamarine3", fill="aquamarine2", bins = 15)+
  geom_density(alpha=0.3, fill="aquamarine1", colour="darkseagreen3") +
  labs(title = "Histograms of Credit/Debit Card Transaction Volume in Turkey over 2015-2020 observed Monthly", 
       x = "Transaction Volume (Thousand TL)",y = "Density") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45,hjust = 0.5)) +
  facet_wrap(~year(Date))
```

  A brief glimpse is enough to say that the mean of transaction volume is increasing every year. This informs the reader about a trend line with positive slope. Also, variance of this measure is increasing every year, a normal-like distribution seems to have tendency to uniform-like distribution. These increments in variance can be caused by different increases in purchasing power in different social economic classes, or they are just a result of increasing variance of transactions. We cannot say anything about seasonality or cyclic structure from these histograms. It will be mentioned with plots of three measures below.
  
## Operation Amount with Credit and Debit Cards
  
  These histograms below include number of operations with credit/debit cards. Note that 2015 has 9 monthly values and 2020 has 10 monthly values. There are 15 bins in the x axis.
  
```{r,echo = FALSE}
ggplot(data = credit_cards, aes(x= credit_operation_amount))+
  geom_histogram(aes(y=..density..), color="lightpink1", fill="lightpink2", bins = 15)+
  geom_density(alpha=0.3, fill="mistyrose2", colour="mistyrose3") +
  labs(title = "Histograms of Credit/Debit Card Operations in Turkey over 2015-2020 observed Monthly", 
       x = "Number of Operations",y = "Density") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45,hjust = 0.5)) +
  facet_wrap(~year(Date))
```
  
  These histograms have similar behavior with the first ones. Not only mean but also variance is increasing over years. The difference between these two histogram sets is that operation amounts have more gaps between values compared to transaction volume. Parts of society may start using cards and cash in a more distinct ways.
  
## Credit Card Usage Index (April 2015 = 100)

  These histograms below include credit card usage index in Turkey. Note that 2015 has 9 monthly values and 2020 has 10 monthly values. There are 10 bins in the x axis.

```{r , echo = FALSE}
ggplot(data = credit_cards, aes(x= credit_index))+
  geom_histogram(aes(y=..density..), color="lightblue1", fill="lightblue2", bins = 10)+
  geom_density(alpha=0.3, fill="lightcyan2", colour="lightcyan3") +
  labs(title = "Histograms of Index of Payments with Cards in Turkey over 2015-2020 observed Monthly", 
       x = "Index (04-2015 = 100)",y = "Density") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45,hjust = 0.5)) +
  facet_wrap(~year(Date))

```
  
  2015 does not give descriptive knowledge due to lack of data. These  histograms of index show that credit card usage has never decreased compared to 04/2015 because all indices are greater than zero. If we look at the mean in 2019, we can say that usage of credit cards is duplicated in 4 years, and continues increasing. However, there is remarkable uniformity in 2020. This has made me think that some people change their payment behavior from cash to card whereas others do not. 
  
# Visual Analysis Combining 3 Measures

## Boxplot of 3 Measures

```{r, echo = FALSE}
credit_cards %>% 
  gather(cards, new_cols , credit_transaction_vol, credit_operation_amount,credit_index ) %>% 
  ggplot(aes(x = Date, y = new_cols , fill = factor(year(Date))))+
  geom_boxplot()+
  scale_fill_brewer(palette = "YlGn", name = "Years") +
  labs(title = "Boxplots of Measures of Credit/Bank Cards Usage 2015-2020", y = "Amount")+
  theme_minimal()+
  facet_grid(rows = vars(cards),scales = 'free')    

```

  These boxplots give general knowledge about dissemination of 3 data sets. These increasing trends and variances over years mean that these 3 measures are highly correlated. Number of operations, transaction volumes, and usage indices have the highest variances respectively. The reader should keep in mind that scales are different for usage index, transaction volume, and number of operations. Indices range between 100-300, but number of operation is in millions, and transaction volumes in billions. For a detailed analysis, we should start examining values in a monthly manner.
  
## Graph for 3 Measures  
  
```{r, echo= FALSE}
cols <- c("Number of Operations" = "orange1", "Transaction Volume" = "tomato","Usage Index" = "firebrick")
ggplot(credit_cards, aes(x = Date)) +
  geom_bar(aes(y= (credit_operation_amount), fill = "Number of Operations"), color = "orange2", stat="identity") +
  geom_bar(aes(y=credit_transaction_vol, fill = "Transaction Volume"), color = "brown2", stat="identity") +
  geom_line(aes(y = credit_index*2000000,color = "Usage Index"),size = 1) + 
  labs(title = "Measures of Credit/Debit Card Usage in Turkey between 2015-2020 Reported Monthly", 
       x = "Date") +
  scale_x_date(date_breaks = "years") +
  scale_y_continuous("Volume (Thousand TL) / Operations", sec.axis = sec_axis(~./2000000, name = "Index")) +
  scale_color_manual(values = cols) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 0.5), legend.position = "right", legend.title = element_blank(),
        axis.line.y.right = element_line(color="firebrick"), axis.text.y.right = element_text(color = "firebrick"), 
        axis.title.y.right = element_text(color="firebrick"))


```
  
  Orange bars stand for number of operations, blue bars stand for transaction volume, and brown line stands for credit card usage indices. Note that y axis in the right is for Index due to its different scale. These three measures went up and down concurrently, implying a high correlation. Also, there is some repeated fluctuations over years, meaning that there is cyclic pattern due to seasonality. Credit / debit card usage measures have seasonal drops in autumn, higher values in summer. Last cycle is different than the previous 5 cycles, having a huge drop caused by the pandemic. However, measures could reach the normally-predicted levels and continue increasing in the long run.

# Correlation Analysis

  Correlations between paired columns can be seen below. First one is the correlation between the transaction volumes and operation amounts. Second one is the correlation between transaction volumes and credit card usage indices. Third one is the correlation between credit card usage indices and operation amounts. These coefficients of the 3 measures are considerably high, reflecting the numerical aspect of our discussion above. 

```{r, echo = FALSE}
print(cor(credit_cards$credit_transaction_vol, credit_cards$credit_operation_amount))
print(cor(credit_cards$credit_transaction_vol, credit_cards$credit_index))
print(cor(credit_cards$credit_index, credit_cards$credit_operation_amount))
```
Are the numbers high enough to say that they are correlated? Thanks to R, we can simply do a correlation test. This test uses Pearson's correlation test by default. Null hypothesis is that there is a significant correlation between our paired variables. Results revealed a significant correlation between our variables, thus we fail to reject our null hypothesis in 95% confidence interval.

```{r, echo = FALSE}

print(cor.test(credit_cards$credit_transaction_vol, credit_cards$credit_operation_amount))
print(cor.test(credit_cards$credit_transaction_vol, credit_cards$credit_index))
print(cor.test(credit_cards$credit_index, credit_cards$credit_operation_amount))

```

## Scatter Plots and Correlation Table

  Scatter plots are useful to see correspondent behavior between two data sets. Distributions in diagonal, correlations in upper triangle, and scatter plots in lower triangle can be seen below. *** means that p-value is low enough for alpha = 0.001 (99.9% confidence interval). Histograms of the variables and Pearson's product-moment correlational tests are mentioned above. However, scatterplots are introduced here.


```{r, echo = FALSE}
ggpairs(credit_cards[,-1]) + theme_minimal() + 
  labs(title ="Scatter Plots and Correlations of Credit/Debit Card Usage in Turkey 2015-2020")+
  theme(axis.text.x = element_text(angle = 45,hjust = 0.5))
```

  All three scatter plots seem to be decent enough to fit in a line. The scatter plot where Credit / debit cards' transaction volumes and operation amounts cross has the best linearity. Also, this pair has the correlation coefficient that is closest to 1. Thus, it can be concluded that increase in the transaction volume is mainly derived from increase in transaction numbers. Other two scatterplots have a slight deviation from a line but their paired correlation coefficients are high enough. This indicates that transaction volume and operation amount have a considerable role in calculating usage indices, but there are some other factors that we can not cover here.
  
  
# Conclusion

  In this study, the main question was "Are there any correlations between credit/ debit cards transaction volume, number of operations, and their usage indices?". We found that transaction volume- number of operations, number of operations - usage index, usage index - transaction volume pairs are highly correlated, and that null hypothesis could not be rejected at 95% significance level. Also, we could grasp that these three measures have trend line with some seasonal fluctuations with their increasing mean and variance. The reader could see the effect of the pandemic in the beginning of 2020. 
  Consequently, changes in the credit/debit cards' transaction volumes covary with the  changes in the number of operations with credit/debit card. Credit card usage index numbers are a result of some complicated calculations, these two variables also covary so we can say that transaction volume and number of operations are important as an input for these calculations. In the linear model,transaction volume and operation amount are considered as independent variables, and credit card usage index as dependent variable. We could check whether our linear model is suitable enough or has significant amount of error by summary function.
  
  
```{r, include = TRUE}
fit <- lm(credit_index~credit_operation_amount + credit_transaction_vol, data = credit_cards)
summary(fit)

```
  P-value of the overall model is neglicable and can be considered as 0. The intercept and our two variables' p-values are also small enough in a %99,9 confidence interval. Adjusted R-squared is a term that shows our model's values fits successfully within the original data. The adjused R-square is approaching to 1, so this linear model could be a good start to make the regression model for these data sets.
  
# Appendix

[Rmd file](\fall20-sudeyagmur\files\hw2\HW2.Rmd)

# References

[EVDS](https://evds2.tcmb.gov.tr/index.php?/evds/serieMarket)
[Stack Overflow](https://stackoverflow.com/)
[Towards Data Science](https://towardsdatascience.com/)
[STHDA](http://www.sthda.com/english/)
[R Documentation](https://www.rdocumentation.org/)






